---
title: "bebida-optimization-result-analysis"
author: "Michael Mercier"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages('rjson')
```

## Experiment Design

Running on Gros cluster on Nancy on 16 nodes (288 cores) with the following heuristics:
- NoHPC: In order to have a control experiment, run the Bebida app without HPC workload
- None: Raw bebida implementation without optimization
- Punch: Create jobs at submission time with arbitrary resource requests
- Annotated: Create jobs on submission using resource annotations to shape the job
- TimeCritical: Use resource quota to dedicate a dynamic set of resources to Bebida application

The Deadline heuristic is similar to Punch Annotated but a deadline provided. It's goal is to run before a given deadline. Because it does not follow the same objective it is not compared to the others.

## Execution time

Initialize Libraries
```{r}
library(tidyverse)
library(viridisLite)
library(rjson)
```

Get data from the experiment result directory:
```{r bdaDf hpcDf metadata}
resultDir = "/home/mmercier/Projects/bebida-optimization-service/result-all-2/"

expeDirs = Sys.glob(sprintf("%s/*", resultDir))
expeDirs
```

Extract BDA applications metrics:
```{r}

bdaDf <- data.frame(heuristic = character(), execution_time = integer())

for (expeDir in expeDirs) {
  execTime = array()
  metadata <- fromJSON(file=sprintf("%s/metadata.json", expeDir))
  
  for (iteration in 1:metadata$nb_app_run) {
    podState <- fromJSON(file=sprintf("%s/spark-app-pi-%d-pod.json", expeDir, iteration))
  
    startTime = parse_datetime(podState$status$containerStatuses[[1]]$state$terminated$startedAt)
    endTime = parse_datetime(podState$status$containerStatuses[[1]]$state$terminated$finishedAt)

    bdaDf <- bdaDf %>% add_row(heuristic = metadata$heuristic, execution_time = as.numeric(difftime(endTime,startTime),units="secs"))
  }
}
bdaDf
```

Extract HPC Jobs metrics.

WARNING: Some job give a submission time after the starting time. which gives a negative waiting time. These jobs are excluded.
```{r}

hpcDf <- data.frame(heuristic = character(), job_id = character(), job_name = character(), waiting_time = integer())

for (expeDir in expeDirs) {

  hpcJobs <- fromJSON(file=sprintf("%s/oar-jobs.json", expeDir))
  metadata <- fromJSON(file=sprintf("%s/metadata.json", expeDir))
  
  for (job in hpcJobs) {
    submitTime = as.POSIXct(job$submission_time, origin="1970-01-01")
    startTime = as.POSIXct(job$start_time, origin="1970-01-01")
    job_id = as.character(job$id)
    waiting_time = as.numeric(difftime(startTime,submitTime),units="secs")

    if (waiting_time > 0 && ! grepl("BEBIDA_NOOP", job$name, fixed = TRUE)) {
      hpcDf <- hpcDf %>% add_row(
        heuristic=metadata$heuristic, job_id=job_id, job_name=job$name, waiting_time=waiting_time) 
    }
  }
}
hpcDf
```

## Plots of BDA app

```{r}
bdaDf %>%
  ggplot() +
  geom_boxplot(aes(x=heuristic, y=execution_time, fill=heuristic, group=heuristic, ymin=0))
```

Transform raw data to stats
```{r}
bdaDfStats <- bdaDf %>%
  group_by(heuristic) %>%
  summarise( 
    n=n(),
    mean=mean(execution_time),
    sd=sd(execution_time)
  ) %>%
  mutate( se=sd/sqrt(n))  %>%
  mutate( ic=se * qt((1-0.05)/2 + .5, n-1))
bdaDfStats
```


```{r}
bdaDfStats %>%
  ggplot() +
  geom_bar( aes(x=heuristic, y=mean, fill=heuristic), stat="identity", alpha=0.5) +
  geom_errorbar(aes(x=heuristic, ymin=mean-sd, ymax=mean+sd), width=0.4, colour="orange", alpha=0.9, size=1.3) +
  ggtitle("Mean execution time in seconds (with standard deviation)")

```
## Plot for HPC

```{r}
hpcDf  %>%
  ggplot() +
  geom_boxplot(aes(x=heuristic, y=waiting_time, fill=heuristic, group=heuristic, ymin=0))
```